---
title: "Case Study C - HIV"
author: "Oumaima Al Qoh, Francisco Arrieta, Lucia Camenisch, Manuela Giansante, Emily Schmidt, Camille Beatrice Valera"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true # creating a table of contents (toc)
    toc_float: 
      collapsed: false # toc does not collapse and is shown as a sidebar (toc_float)
    number_sections: true # document sections are numbered
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

```{r }
# clear the database
rm(list=ls()) 
graphics.off()
shell("cls")
```

```{r packages}
library(data.table)
library(ggplot2)
library(e1071)
library(missForest)
library(caret)
library(MASS)
library(gridExtra)
```

```{r}
aids <- fread("Caids.csv")

aids[, sex := as.factor(sex)]
aids[, type := as.factor(type)]

new_DF <- aids[rowSums(is.na(aids)) > 0,]
```

```{r}
par(mfrow = c(3,2))

hist(aids$cd4)
hist(sqrt(aids$cd4))
hist(aids$cd8)
hist(sqrt(aids$cd8))
hist(aids$rna)
hist(log(aids$rna))
```

```{r}
aids_melt <- melt(aids[, -1], id.vars = "type")
ggplot(data = aids_melt[type %in% "DP",], aes(value)) + geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "DP")
ggplot(data = aids_melt[type %in% "CP",], aes(value)) + geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  labs(title = "CP")
```

```{r}
cor(aids[, 3:5], use = "complete.obs")
```

```{r}
skewness(sqrt(aids$cd4), na.rm = TRUE)
skewness(sqrt(aids$cd8), na.rm = TRUE)
skewness(log(aids$rna), na.rm = TRUE)
```

```{r}
aids[, sex := ifelse(sex == "m", 0, 1)]
aids[, type := ifelse(type == "DP", 0, 1)] # DP is 1

plot(aids[, 3:5], col = aids$type)

summary(aids)
```

```{r}
aids_completed <- missForest(aids, variablewise = TRUE)

summary(aids)
summary(aids_completed$ximp)
```

```{r}
aids <- aids_completed$ximp
```

## Trees: Data Partitioning

```{r}
library(rpart) # Used for building classification and regression trees
library(RColorBrewer)
library(rattle)  # Graphical Data Interface
library(rpart.plot) # Automatically scales and adjusts the displayed tree for best fit
library(gains)
library(adabag)
library(randomForest)
library(lift) # TopDecileLift()
library(caret)
library(kableExtra)
library(vip)
library(tidyverse) # Data manipulation
library(xgboost) # Fitting the xgboost model
library(pROC)
library(ipred)       #for fitting bagged decision trees

aids[, sex := as.factor(sex)]
aids[, type := as.factor(type)]

# Set the seed for the random number generator for reproducing the partition.
set.seed(1)

# Partitioning into training (60%) and validation (40%) 
train.index <- sample(c(1:dim(aids)[1]), dim(aids)[1]*0.6)
test.index <- setdiff(c(1:dim(aids)[1]), train.index)  

# Collect all the columns with training rows into training set 
train <- aids[train.index, ]
test <- aids[test.index, ]
```

```{r}
prop.table(table(train$type))
prop.table(table(test$type))
```

-   77.1% of the training data is type 0, DP\
-   22.9% of the training data is type 1, CP\
-   74.1% of the test data is type 0, DP\
-   25.9% of the test data is type 1, CP\

## Deep Classification Decision Tree

```{r Deep Tree}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

# Create deep tree (page 219)
deep_class_tree <- rpart(formula = type ~ ., data = train, method = "class", cp = 0, minsplit = 2)

# Generates a cost complexity parameter table that provides the complexity parameter value
#summary(deep_class_tree)

# Plot tree
prp(deep_class_tree, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10, box.col  = ifelse(deep_class_tree$frame$var == "<leaf>", 'lightblue', 'white'))

# Count number of leaves 
length(deep_class_tree$frame$var[deep_class_tree$frame$var == "<leaf>"]) 

# Variable importance, deep tree
kable_styling(kable(round(100*deep_class_tree$variable.importance / sum(deep_class_tree$variable.importance), 2), col.names = "Importance %"), full_width = TRUE) |>  kable_classic() 

printcp(deep_class_tree)

#Get the lowest CP value from CP table
deep_class_tree$cptable[which.min(deep_class_tree$cptable[,"xerror"]),"CP"]
```

-   Fit a binary model, type (0,1)\
-   cp = 0: Complexity parameter is a tuning parameter that *controls the amount of pruning applied*. There is no pruning done in a deep tree.\
-   rpart(): Builds the tree recursively by splitting the data into smaller subsets based on the values of the predictor variables. However, this can lead to overfitting - the tree becomes too complex ad fits the noise in the data rather than the underlying pattern.\
-   To avoid overfitting, one can prune the tree by removing splits or branches from the tree.\
-   A larger cp: More aggressive the pruning, simplifying the tree and reduces the possibility of overfitting the model.\

-   Deep tree has a total of 37 terminal nodes.\

Variance importance measures the contribution of each predictor variable in the model of the response variable, `type`. It is computed based on the number of times that predictor is used to split the data in the decision tree as well as the improvement of the model's performance. `cd4` appears to have the biggest influence followed by `cd8` and `rna`. `sex` is much lower than all the other variables.\

Lowest xerror and xstd is nsplit = 2 with a CP of 0.04. All four predictors were used.\

The resulting min.xerror variable of 0.04 contains the value of the complexity parameter associated with the *minimum cross-validated error*. This value can be used to prune the decision tree to the appropriate level of complexity to avoid overfitting.\

```{r}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

#Plot the optimal Cp value
plotcp(deep_class_tree)
```

Plotcp() provides a graphical representation to the cross validated error summary. The cp values are plotted against the geometric mean to depict the deviation until the minimum value is reached.\
- The x-val relative error (cross-validated error rate) is minimized at size of tree 3 with the cp equaling 0.046.\
- When the "x-val relative error" values start to increase rapidly, it indicates that the decision tree model is starting to overfit the training data and may not generalize well to new, unseen data.\

## Deep Tree Confusion Matrix
```{r Deep Tree Confusion Matrix}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

predict_deep_class_tree <- predict(deep_class_tree, test, type = "class")

# Create confusion matrix for classification tree
conf_mat_tree1  <- confusionMatrix(predict_deep_class_tree, as.factor(test$type))
conf_mat_tree1

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_deep_class_tree, as.factor(test$type))$table)
```

-   Upper left: True Negative - You predicted negative and it's true.\
-   Lower left: False Negative - You predicted negative and it's false, leading to a Type II error\
-   Upper right: False Positive - You predicted positive and it's false, which is a Type I error\
-   Lower right: True Positive - You predicted positive and it's true.\

Recall: From all positive classes, how many we predicted correctly. This value should be as high as possible.\
$$Recall =  \frac{TP}{TP + FN}$$

```{r Deep Tree Recall}
deep_recall1 <- conf_mat_tree1$byClass[6]
deep_recall1
```

The recall score is 78.3%.\

Precision: From all the classes we have predicted as positive, how many are actually positive. This value should be as high as possible.\
$$Recall =  \frac{TP}{TP + FP}$$

```{r Deep Tree Precision}
deep_precision1 <- conf_mat_tree1$byClass[5]
deep_precision1
```

The precision score is 73.8%.\

```{r Acc Spec Spec Deep Tree}
# Accuracy and specificity for tree
tree_res1 = c(conf_mat_tree1$overall[1], conf_mat_tree1$byClass[1], conf_mat_tree1$byClass[2])

tree_res1
```

Accuracy: From all the classes (positive and negative), how many of them we have predicted correctly. This value should be as high as possible.\
Sensitivity: The proportion of true positives among all actual positives. In other words, sensitivity measures how well the model identifies the positive class. Higher the value, the better!\
$$ Sensitivty = TP/(TP+FN)$$

Specificity: The proportion of true negatives among all actual negatives. In other words, specificity measures how well the model identifies the negative class. Higher the value, the better!\
$$ Specificity = TN/(TN+FP)$$

The accuracy s 63.3%, while Sensitivity is 78.3% and Specificity is 20.7%. This model does not predict the negative class as well.\

F-measure: It is difficult to compare two models with low precision and high recall or vice versa. So to make them comparable, we use F-Score. F-score helps to measure Recall and Precision at the same time. It uses Harmonic Mean in place of Arithmetic Mean by punishing the extreme values more.\
$$F-measure =  \frac{2*Recall*Precision}{Recall + Precision}$$

```{r Deep Tree F-measure}
(2*deep_recall1 * deep_precision1)/(deep_recall1 + deep_precision1)
```

The f-measure score is 76.2%.

```{r}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

#Obtaining predicted probabilites for Test data
deep.probs = predict(deep_class_tree,
                 newdata = test,
                 type = "prob")
head(deep.probs)

# Calculate ROC curve
rocCurve.deep <- roc(test$type,deep.probs[,"0"])

# Plot the ROC curve
plot(rocCurve.deep,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.deep)
```

AUC is 50.5%.

After computing the optimal xerror and CP, we now prune the tree back to get the best model.

```{r}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

prune_tree <- prune(deep_class_tree, cp = 0.026316)

fancyRpartPlot(prune_tree, uniform = TRUE, caption = NULL, palettes = "GnBu", digits = -3, main = "Pruned Classification Tree")
```

-   It appears that the optimal tree only has one split at `rna` \< 274,063, known as the root node.\
-   This predictor value is used to split the data into two groups based on the threshold of 274,063.
-   Observations on the left satisfy the condition that `rna` is less than that value. Therefore, they are classified as DP (0), otherwise CP (1) on the right side which would be equivalent to stating the the remaining observations are equal or greater than 274K.\
-   There are three terminal nodes compared to the 38 that was seen within the deep classification model.\
-   Five nodes of a decision tree, In each node, there is either a 0/1 which represents DP or CP, and two decimals within. The 0 or 1 indicates that activity that the group wants to take. The number on the left shows the probability of people in this group to take this activity. While the right-hand probability is the opposite, they do not take the activity.\
-   In the root node, we see a 0 which represents the DP, 77.1% which is the probability that only one partner is HIV-positive, and only 22.9% of those individuals are both HIV-positive. Depending on if the respectie observation meets the criteria plays out how the terminal nodes read. There are 83.7% of DP that have an `rna` less than the condition. Within that left node, there is another proportion which is 16.3%, representing those that do not meet that criteria. Together, they equal 100%. In the other node, once again, there are two values that are represented, and can be interpreted in the same manner.\

```{r eval=FALSE, include=FALSE}
# Variable importance, prune tree
kable_styling(kable(round(100*prune_tree$variable.importance / sum(prune_tree$variable.importance), 2), col.names = "Importance %"), full_width = TRUE) |>  kable_classic() 
```

-   In the deep tree, all predictor values were used. This model no longer uses `sex`, and this was intuitive since in the previous model, it only had an importance of approximately 4%. `cd8` used to be the most important factor, but now `rna` takes the lead, followed by the other two cells.\

```{r eval=FALSE, include=FALSE}
printcp(prune_tree)
```

This verifies the optimal choice of 2-split.\

## Pruned Tree Confusion Matrix
```{r Pruned Tree Confusion Matrix, eval=FALSE, include=FALSE}
predict_prune_tree <- predict(prune_tree, test, type = "class")

# Create confusion matrix for classification tree
conf_mat_tree2  <- confusionMatrix(predict_prune_tree, as.factor(test$type))
conf_mat_tree2

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_prune_tree, as.factor(test$type))$table)
```

```{r Pruned Tree Recall, eval=FALSE, include=FALSE}
deep_recall2 <- conf_mat_tree2$byClass[6]
deep_recall2
```

The recall score is 90.4%.

```{r Pruned Tree Precision, eval=FALSE, include=FALSE}
deep_precision2 <- conf_mat_tree2$byClass[5]
deep_precision2
```

The precision score is 80.6%.

```{r Acc Spec Spec Pruned Tree ,eval=FALSE, include=FALSE}
# Accuracy and specificity for tree
tree_res2 = c(conf_mat_tree2$overall[1], conf_mat_tree2$byClass[1], conf_mat_tree2$byClass[2])
tree_res2
```

The accuracy s 76.8%, while Sensitivity is 90.3% and Specificity is 37.9%. This model continues to not predict the negative class as well as it has decreased since the deep tree model. Both the accuracy and sensitivity have increased though.\

```{r Pruned Tree F-measure, eval=FALSE, include=FALSE}
(2*deep_recall2 * deep_precision2)/(deep_recall2 + deep_precision2)
```

The f-measure score is 85.2%.

```{r eval=FALSE, include=FALSE}
#Obtaining predicted probabilites for Test data
prune.probs = predict(prune_tree,
                 newdata = test,
                 type = "prob")
head(prune.probs)

# Calculate ROC curve
rocCurve.prune <- roc(test$type,prune.probs[,"0"])

# Plot the ROC curve
plot(rocCurve.prune,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.prune)
```

AUC is 35.85%.

This next part could be for general knowledge?
## Classification Decision Tree

```{r}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

# Fit the model
class_tree <- rpart(type ~ ., data = train, method = "class") # cp default is 0.01

# Plot the classification tree
fancyRpartPlot(class_tree, caption = NULL, main = "Classification Tree", palettes = "GnBu", digits = -3)
```

-   The classification tree is smaller than the deep tree but approximately the same as the pruned tree. This in due to the CP being set to the default at 0.01. The deep tree has a value of 0, while the pruned tree's CP was higher at 0.05, therefore, causing the tree to be more pruned.

```{r}
# Variable importance, deep tree
kable_styling(kable(round(100*class_tree$variable.importance / sum(class_tree$variable.importance), 2), col.names = "Importance %"), full_width = TRUE) |>  kable_classic() 
```

-   The variable importance is a mixture of the deep tree and pruned results. This model uses all predictors, but in order of the pruned tree with `rna` being the most relevant. Once again, `sex` is barely used for splitting and is pretty much irrelevant.


```{r Deep Tree Confusion Matrix}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

predict_class_tree <- predict(class_tree, test, type = "class")

# Create confusion matrix for classification tree
conf_mat_tree1  <- confusionMatrix(predict_class_tree, as.factor(test$type))
conf_mat_tree1

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_deep_class_tree, as.factor(test$type))$table)
```

```{r}
printcp(class_tree)
```

Right away, we see that the optimal choice for xerror and xstd are at split 2, which corresponds to a CP or .010. This is lower than the the deep tree but similar to the pruned tree. Therefore, will explore that option now.

```{r}
prune_tree_2 <- prune(class_tree, cp = 0.010000)

fancyRpartPlot(prune_tree_2, uniform = TRUE, caption = NULL, palettes = "GnBu", digits = -3, main = "Pruned Classification Tree")
```

Once we recompute the pruned classification tree, we get the same exact result from what was concluded with the pruned tree. Therefore, the conclusion for the classification decision tree is that two splits on `rna` 274,063, and cd8 >- 626 will provide the most optimal predictions on the test data set. 

**UNSURE IF THIS PART IS NECESSARY!!! LETS DISCUSS**
```{r eval=FALSE, include=FALSE}
par(mfrow=c(1,2)) # Set two plots side-by-side

# Predict probabilities for the classification tree with the validation set
pred_prob = predict(prune_tree, test, type = "prob")

# Create dataframe with the actual and probability values for the validation set
ClassTreeDF = data.frame(actual = test$type, prob = predict(prune_tree, test, type = "prob")[,2])

# Compute gains for lift chart (page 138)
gain <- gains(predicted = ClassTreeDF$prob, actual = as.numeric(as.character(ClassTreeDF$actual)), groups=dim(ClassTreeDF)[1], ties.method = c("first"))

# Plot the lift chart (page 138)
plot(c(0, gain$cume.pct.of.total*sum(as.numeric(as.character(ClassTreeDF$actual)))) ~ c(0, gain$cume.obs), main = "Gains Chart",
xlab = "# Cases", ylab = "Cumulative", type="l")

# Plot the lift chart diagonal (page 138)
lines(c(0,sum(as.numeric(as.character(ClassTreeDF$actual)))) ~ c(0,dim(ClassTreeDF)[1]), lty = 2, col = "blue")

# Compute deciles (page 249)
gain.decile <- gains(predicted = ClassTreeDF$prob, actual = as.numeric(as.character(ClassTreeDF$actual)),ties.method = c("first"))

# Plot decile-wise chart (page 249)
barplot(gain.decile$mean.resp / mean(as.numeric(as.character(ClassTreeDF$actual))), names.arg = gain.decile$depth, main = "Decile-wise Lift Chart", xlab = "Percentile", ylab = "Mean Response")
```

```{r eval=FALSE, include=FALSE}
# Calculates the top-decile lift which expresses the incidence in 10% of observations 
TopDecileLift(pred_prob[,2], test$type)
```

**COMMENTS FROM ANOTHER ASSIGNMENT, PLEASE IGNORE WRONG VALUES** Some useful tools for assessing model classification are the gains chart and decile-wise lift chart. The "lift" over the base curve indicates for a given number of cases, the additional res ponders that you can identify by using the model. This helps derive good accuracy measures. The decile-wise lift chart takes 10% of the records that are ranked by the model as "most probable 1's" yields 33 times as many 1's as would simply selecting 10% of the records at random. The lift will vary with the number of records we choose to act on. A good classifier will give us a high lift when we act on only a few records.

The TopDecileLift() function allows a user to calculate the top-decile lift, a metric that expresses how the incidence in the 10% customers with the highest model predictions compares to the overall sample incidence. A lift of 1 is expected for a random model. For this classification tree though, a lift of 1.914 indicates that in the 10% highest predictions, roughly two times more positive cases are identified by the model than would be expected for a random selection of instances.

```{r eval=FALSE, include=FALSE}
# Produce table with gains
gain_decile <- gains(actual = as.numeric(as.character(test$type)), 
                     predicted = predict(prune_tree, test, type = "prob")[,2])
gain_decile
```

```{r eval=FALSE, include=FALSE}
# Plot decile chart
## Height of bars
heights <- gain_decile$mean.resp/mean(as.numeric(as.character(test$type)))

midpoints <- barplot(heights, names.arg = gain_decile$depth, ylim = c(0,2.2),
                     xlab = "Percentile", ylab = "Mean Response", 
                     main = "Decile-wise chart for validation data, tree model")

text(x=midpoints, y = heights, labels = round(heights, 1), pos = 3.5, cex = 0.7, col = "red")
```

## Boosting
Resource: https://www.projectpro.io/recipes/tune-hyper-parameters-grid-search-r
```{r}
# Specifying the CV technique which will be passed into the train() function later and number parameter is the "k" in K-fold cross validation
train_control = trainControl(method = "cv", number = 5, search = "grid")

set.seed(12)
# Customizing the tuning grid
gbmGrid <-  expand.grid(max_depth = c(3, 5, 7), 
                        nrounds = (1:10)*50, # number of trees
                        # default values below
                        eta = 0.3,
                        gamma = 0,
                        subsample = 1,
                        min_child_weight = 1,
                        colsample_bytree = 0.6)

# Training a XGboost classification tree model while tuning parameters
model = train(type~., data = train, method = "xgbTree", trControl = train_control, tuneGrid = gbmGrid)

# Summarizing the results
print(model)
```
When we train a model, the best parameters are determined for each independent variable. For example, in linear reggression modelling, the coefficients of each independent variable is considered as a parameter i.e. they are found during the training process.\

On the hand, Hyperparameters are are set by the user before training and are independent of the training process. For example, depth of a Decision Tree. These hyper parameters affects the performance as well as the parameters of the model. Hence, they need to be optimised. There are two ways to carry out Hyperparameter tuning:

- Grid Search: This technique generates evenly spaced values for each hyperparameters and then uses Cross validation to find the optimum values.\
- Random Search: This technique generates random values for each hyperparameter being tested and then uses Cross validation to find the optimum values.
In this recipe, we will discuss how to build and optimise size of the tree in XGBoost using hyperparameter tuning using Grid Search.\

Recently, researchers and enthusiasts have started using ensemble techniques like XGBoost to win data science competitions and hackathons. It outperforms algorithms such as Random Forest and Gadient Boosting in terms of speed as well as accuracy when performed on structured data.\

XGBoost uses ensemble model which is based on Decision tree. A simple decision tree is considered to be a weak learner. The algorithm build sequential decision trees were each tree corrects the error occuring in the previous one until a condition is met.\

MODEL RESULTS
- Tuning parameter 'eta' was held constant at a value of 0.3.\
- Tuning parameter 'gamma' was held constant at a value of 0.\
- Tuning parameter 'colsample_bytree' was held constant at a value of 0.6.\
- Tuning parameter 'min_child_weight' was held  constant at a value of 1.\
- Tuning parameter 'subsample' was held constant at a value of 1.\
- Accuracy was used to select the optimal model using the largest value.\
- The final values used for the model were nrounds = 50, max_depth = 5, eta = 0.3, gamma = 0, colsample_bytree = 0.6, min_child_weight = 1 and subsample = 1.\ 

Used 166 samples and 4 predictors. There was no pre-processing and the resampling used a cross-validation of 5 fold. The results across the tuning parameters turned out to have the highest accuracy at 77.1%, which had a max_depth of 5, nrounds of 50 (number of trees), and $\kappa$ of .218.

## Boosting Confusion Matrix
```{r Boosting Tree Confusion Matrix}
set.seed(12)

predict_boosting <- predict(model, test)

# Create confusion matrix for classification tree
conf_mat_tree3  <- confusionMatrix(predict_boosting, as.factor(test$type))
conf_mat_tree3

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_boosting, as.factor(test$type))$table)
```

```{r}
set.seed(12)

vip::vip(model, aesthetics = list(fill = "lightblue")) +
  ggtitle("Boosted Tree Variable Importance") + # Title name
  xlab("Variable") + # Label names
  theme_classic() # A classic theme, with x and y axis lines and no grid lines
```

There are only three predictors that are used within the model, `sex` being eliminated. `rna` is seen as the most important variable as it was in the pruned tree.

```{r Boost Recall}
deep_recall3 <- conf_mat_tree3$byClass[6]
deep_recall3
```

The recall score is 80.7%.

```{r Boost Precision}
deep_precision3 <- conf_mat_tree3$byClass[5]
deep_precision3
```

The precision score is 77.9%.

```{r Acc Spec Spec Boost}
# Accuracy and specificity for tree
tree_res3 = c(conf_mat_tree3$overall[1], conf_mat_tree3$byClass[1], conf_mat_tree3$byClass[2])
tree_res3
```

The accuracy s 68.8%, while Sensitivity is 80.7% and Specificity is 34.5%. This model also does not predict the negative class  well as it has decreased since the pruned tree model.\

```{r Boost F-measure}
(2*deep_recall3 * deep_precision3)/(deep_recall3 + deep_precision3)
```

The f-measure score is 79.3%.

```{r}
#Obtaining predicted probabilites for Test data
boost.probs = predict(model,
                 newdata = test,
                 type = "prob")
head(boost.probs)

# Calculate ROC curve
rocCurve.boost <- roc(test$type,boost.probs[,"0"])

# Plot the ROC curve
plot(rocCurve.boost,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.boost)
```

AUC is 62.61%.

**UNSURE IF THIS PART IS NECESSARY!!! LETS DISCUSS**
```{r eval=FALSE, include=FALSE}
par(mfrow=c(1,2)) # Set two plots side-by-side

# Create dataframe with the actual and probability values for the validation set
Boost_Tree = data.frame(actual = valid$type, prob = boost_predict$prob[,2])

# Compute gains for lift chart (page 138)
gain <- gains(predicted = Boost_Tree$prob, actual = as.numeric(as.character(Boost_Tree$actual)), groups=dim(Boost_Tree)[1])

# Plot the lift chart (page 138)
plot(c(0, gain$cume.pct.of.total*sum(as.numeric(as.character(Boost_Tree$actual)))) ~ c(0, gain$cume.obs), main = "Gains Chart",
xlab = "# Cases", ylab = "Cumulative", type="l")

# Plot the lift chart diagonal (page 138)
lines(c(0,sum(as.numeric(as.character(Boost_Tree$actual)))) ~ c(0,dim(Boost_Tree)[1]), lty = 2, col = "blue")

# Compute deciles (page 249)
gain.decile <- gains(predicted = Boost_Tree$prob, actual = as.numeric(as.character(Boost_Tree$actual)),ties.method = c("first"))

# Plot decile-wise chart (page 249)
barplot(gain.decile$mean.resp / mean(as.numeric(as.character(Boost_Tree$actual))), names.arg = gain.decile$depth, main = "Decile-wise Lift Chart", xlab = "Percentile", ylab = "Mean Response")
```

```{r eval=FALSE, include=FALSE}
# Calculates the top-decile lift which expresses the incidence in 10% of observations 
TopDecileLift(boost_predict$prob[,2], valid$type)
```

**COMMENTS FROM ANOTHER ASSIGNMENT, PLEASE IGNORE WRONG VALUES** 
For this model, it has a lift of 1.939. The value for boosted trees is slightly higher than Classification Trees which shows how positive cases are identified by the model than would be expected for a random selection of instances.

## Bagging
```{r}
# Code for running bagging tree 
#bagged <- bagging(type ~ ., data = train, coob = TRUE)
bag <- bagging(formula = type ~ ., data = train, nbagg = 150, coob = TRUE, control = rpart.control(minsplit = 2, cp = 0))
bag
```

- Note that we chose to use 150 bootstrapped samples to build the bagged model and we specified coob to be TRUE to obtain the estimated out-of-bag error.\
- We also used the following specifications in the rpart.control() function:
- minsplit = 2: This tells the model to only require 2 observations in a node to split.\
cp = 0. This is the complexity parameter. By setting it to 0, we don’t require the model to be able to improve the overall fit by any amount in order to perform a split.\

## Bagging Confusion Matrix
```{r Bagging Tree Confusion Matrix}
predict_bagged <- predict(bag, test)

# Create confusion matrix for classification tree
conf_mat_tree4 <- confusionMatrix(predict_bagged, test$type)
conf_mat_tree4

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_bagged, as.factor(test$type))$table)
```

```{r}
#calculate variable importance
VI <- data.frame(var=names(train[,-4]), imp=varImp(bag))

#sort variable importance descending
VI_plot <- VI[order(VI$Overall, decreasing=TRUE),]

#visualize variable importance with horizontal bar plot
data<-data.table(name=row.names(VI),value=VI$Overall)

data[,ggplot(.SD, aes(x=reorder(name, VI$Overall), y=VI$Overall)) + 
  geom_bar(stat = "identity", fill = "lightblue") +
  xlab("Variable") + # Label names + 
  ylab("Importance") + # Label names + 
  ggtitle("Bagged Tree Variable Importance") + # Title name
  theme_classic() + # A classic theme, with x and y axis lines and no grid lines
  coord_flip(),]
```

All predictors are used within the model. `cd8` is back at being the most important, with `rna` and `cd4` being approximately as important as `sex` continues to trail behind.

```{r Bagging Recall}
deep_recall4 <- conf_mat_tree4$byClass[6]
deep_recall4
```

EDIT VALUE: The recall score is 88.0%.

```{r Bagging Precision}
deep_precision4 <- conf_mat_tree4$byClass[5]
deep_precision4
```

EDIT VALUE: The precision score is 79.3%.

```{r Acc Spec Spec Bagging}
# Accuracy and specificity for tree
tree_res4 = c(conf_mat_tree4$overall[1], conf_mat_tree4$byClass[1], conf_mat_tree4$byClass[2])
tree_res4
```

The accuracy s 74.1%, while Sensitivity is 88.0% and Specificity is 34.5%. This model also does not predict the negative class  well as it has decreased since the pruned tree model.\

```{r Bagging F-measure}
(2*deep_recall4 * deep_precision4)/(deep_recall4 + deep_precision4)
```

EDIT VALUE: The f-measure score is 79.3%.

```{r}
#Obtaining predicted probabilites for Test data
bag.prob = predict(bag,
                 newdata = test,
                 type = "prob")
head(bag.prob)

# Calculate ROC curve
rocCurve.bag <- roc(test$type,bag.prob[,"0"])

# Plot the ROC curve
plot(rocCurve.bag,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.bag)
```

AUC is 60.01%.

**UNSURE IF THIS PART IS NECESSARY!!! LETS DISCUSS**
```{r eval=FALSE, include=FALSE}
par(mfrow=c(1,2)) # Set two plots side-by-side

# Create dataframe with the actual and probability values for the validation set
Bagged_Tree = data.frame(actual = valid$type, prob = pred_bagged$prob[,2])

# Compute gains for lift chart (page 138)
gain <- gains(predicted = Bagged_Tree$prob, actual = as.numeric(as.character(Bagged_Tree$actual)), groups=dim(Bagged_Tree)[1])

# Plot the lift chart (page 138)
plot(c(0, gain$cume.pct.of.total*sum(as.numeric(as.character(Bagged_Tree$actual)))) ~ c(0, gain$cume.obs), main = "Gains Chart",
xlab = "# Cases", ylab = "Cumulative", type="l")

# Plot the lift chart diagonal (page 138)
lines(c(0,sum(as.numeric(as.character(Bagged_Tree$actual)))) ~ c(0,dim(Bagged_Tree)[1]), lty = 2, col = "blue")

# Compute deciles (page 249)
gain.decile <- gains(predicted = Bagged_Tree$prob, actual = as.numeric(as.character(Bagged_Tree$actual)),ties.method = c("first"))

# Plot decile-wise chart (page 249)
barplot(gain.decile$mean.resp / mean(as.numeric(as.character(Bagged_Tree$actual))), names.arg = gain.decile$depth, main = "Decile-wise Lift Chart", xlab = "Percentile", ylab = "Mean Response")
```

```{r eval=FALSE, include=FALSE}
# Calculates the top-decile lift which expresses the incidence in 10% of observations 
TopDecileLift(pred_bagged$prob[,2], valid$type)
```

**COMMENTS FROM ANOTHER ASSIGNMENT, PLEASE IGNORE WRONG VALUES** 
For this model, it has a lift of 1.939. The value for bagged trees is slightly higher than Classification Trees, but has the same lift as the boosted tree. This shows how positive cases are identified by the model than would be expected for a random selection of instances.

## Random Forest
```{r}
# Code for running random forest
random_forest <- randomForest(train$type ~ ., data = train)
random_forest
```

- Number of trees: 500\
- Number of variables tried at each split: 2\
- OOB: 24.1%\

```{r}
# Number of trees with lowest MSE
# Compared with 80 and 100 trees, the trade-off with both of those were small so we stuck with 60 trees
which.min(random_forest$mse)
```

**UNSURE WHY THIS IS SHOWING AS 0**

## Random Forest Confusion Matrix
```{r}
# Predict for validation data based off model (page 231)
pred_RF <- predict(random_forest, test)

# Create confusion matrix for classification tree
conf_mat_tree5  <- confusionMatrix(as.factor(pred_RF), test$type, positive = "0")
conf_mat_tree5

# Create fourfoldplot
fourfoldplot(confusionMatrix(pred_RF, as.factor(test$type))$table)
```

```{r}
options(scipen = 9999)

# Create variable importance chart
vip::vip(random_forest, aesthetics = list(fill = "lightblue")) +
  ggtitle("RandomTree Variable Importance") + # Title name
  xlab("Variable") + # Label names
  theme_classic() # A classic theme, with x and y axis lines and no grid lines
```

All predictors are used, and follow the same order as the classification model with 8 terminal modes.  `rna` does not continue to be the most important variable as `cd8` is approximately as relevant in this model.

```{r RF Recall}
deep_recall5 <- conf_mat_tree5$byClass[6]
deep_recall5
```

The recall score is 88.0%.

```{r RF Precision}
deep_precision5 <- conf_mat_tree5$byClass[5]
deep_precision5
```

The precision score is 79.3%.

```{r Acc Spec Spec RF}
# Accuracy and specificity for tree
tree_res5 = c(conf_mat_tree5$overall[1], conf_mat_tree5$byClass[1], conf_mat_tree5$byClass[2])
tree_res5
```

The accuracy s 74.1%, while Sensitivity is 88.0% and Specificity is 34.5%. This model also does not predict the negative class  well as it has decreased since the pruned tree model.\

```{r RF F-measure}
(2*deep_recall5 * deep_precision5)/(deep_recall5 + deep_precision5)
```

The f-measure score is 83.4%.

```{r}
#Obtaining predicted probabilites for Test data
RF.probs = predict(random_forest,
                 newdata = test,
                 type = "prob")
head(RF.probs)

# Calculate ROC curve
rocCurve.rf <- roc(test$type,RF.probs[,"0"])

# Plot the ROC curve
plot(rocCurve.rf,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.rf)
```

AUC is 63.81%.

**UNSURE IF THIS PART IS NECESSARY!!! LETS DISCUSS**
```{r eval=FALSE, include=FALSE}
par(mfrow=c(1,2)) # Set two plots side-by-side

# Predict probabilities for the bagged tree with the validation set
pred_RF <- predict(RF, valid, type = "prob")

# Create dataframe with the actual and probability values for the validation set
RF_Tree = data.frame(actual = valid$type, prob = pred_RF[,2])

# Compute gains for lift chart (page 138)
gain <- gains(predicted = RF_Tree$prob, actual = as.numeric(as.character(RF_Tree$actual)), groups=dim(RF_Tree)[1])

# Plot the lift chart (page 138)
plot(c(0, gain$cume.pct.of.total*sum(as.numeric(as.character(RF_Tree$actual)))) ~ c(0, gain$cume.obs), main = "Gains Chart",
xlab = "# Cases", ylab = "Cumulative", type="l")

# Plot the lift chart diagonal (page 138)
lines(c(0,sum(as.numeric(as.character(RF_Tree$actual)))) ~ c(0,dim(RF_Tree)[1]), lty = 2, col = "blue")

# Compute deciles (page 249)
gain.decile <- gains(predicted = RF_Tree$prob, actual = as.numeric(as.character(RF_Tree$actual)),ties.method = c("first"))

# Plot decile-wise chart (page 249)
barplot(gain.decile$mean.resp / mean(as.numeric(as.character(RF_Tree$actual))), names.arg = gain.decile$depth, main = "Decile-wise Lift Chart", xlab = "Percentile", ylab = "Mean Response")
```

```{r eval=FALSE, include=FALSE}
# Calculates the top-decile lift which expresses the incidence in 10% of observations 
TopDecileLift(pred_RF[,2], valid$type)
```


## Random Forest, Stochastic Gradient Boosting 
```{r}
modelLookup("gbm")

cvcontrol <- trainControl(method = "repeatedcv", number = 5, allowParallel = TRUE)

gbm <- train(type ~ ., data = train, method = "gbm", verbose = F, trControl = cvcontrol)
gbm
```

Used 166 samples and 4 predictors with a cross-validation of 5 fold repeated once. The results across the tuning parameters turned out to have the highest accuracy at 77.7%, which had a interaction_depth of 3, nrounds of 50 (number of trees), and $\kappa$ of .262.

## Random Forest SGB Confusion Matrix
```{r}
# Predict for validation data based off model (page 231)
pred_RFSGB <- predict(gbm, test)

# Create confusion matrix for classification tree
conf_mat_tree6  <- confusionMatrix(pred_RFSGB, test$type)
# I think this line below can be deleted
# conf_mat_tree5 <- confusionMatrix(train$ype, gbm)
conf_mat_tree6

# Create fourfoldplot
fourfoldplot(confusionMatrix(pred_RFSGB, as.factor(test$type))$table)
```


```{r}
options(scipen = 9999)

# Create variable importance chart
vip::vip(random_forest, aesthetics = list(fill = "lightblue")) +
  ggtitle("RandomTree Variable Importance") + # Title name
  xlab("Variable") + # Label names
  theme_classic() # A classic theme, with x and y axis lines and no grid lines
```

All predictors are used, and follow the same order as the classification model with 8 terminal modes.  `rna` does not continue to be the most important variable as `cd8` is approximately as relevant in this model.

```{r RFSBG Recall}
deep_recall6 <- conf_mat_tree6$byClass[6]
deep_recall6
```

EDIT VALUE: The recall score is 88.0%.

```{r RFSBG Precision}
deep_precision5 <- conf_mat_tree5$byClass[5]
deep_precision5
```

EDIT VALUE:The precision score is 79.3%.

```{r Acc Spec Spec RFSBG}
# Accuracy and specificity for tree
tree_res5 = c(conf_mat_tree5$overall[1], conf_mat_tree5$byClass[1], conf_mat_tree5$byClass[2])
tree_res5
```

EDIT VALUES: The accuracy s 74.1%, while Sensitivity is 88.0% and Specificity is 34.5%. This model also does not predict the negative class  well as it has decreased since the pruned tree model.\

```{r RFSBG F-measure}
(2*deep_recall5 * deep_precision5)/(deep_recall5 + deep_precision5)
```

EDIT VALUE: The f-measure score is 83.4%.

```{r}
#Obtaining predicted probabilites for Test data
RF.probs = predict(random_forest,
                 newdata = test,
                 type = "prob")
head(RF.probs)

# Calculate ROC curve
rocCurve.rf <- roc(test$type,RF.probs[,"0"])

# Plot the ROC curve
plot(rocCurve.rf,col=c(1))

# Calculate the area under curve (bigger is better)
auc(rocCurve.rf)
```

EDIT VALUE: The AUC is 00.0%


## Conclusion
```{r}
Model <- c("Deep Tree", "Pruned Tree", "Boosting", "Bagging", "Random Forest", "Random Forest SGB")
Recall <- c(78.3, 90.4, 80.7, 1, 88.0, 1)
Precision <- c(80.2, 80.6, 77.9, 1, 79.3, 1)
Accuracy <- c(69.6, 76.8, 68.8, 1, 74.1, 1)
Sensitivity <- c(78.3, 90.3, 80.7, 1, 88.0, 1)
Specificity <- c(44.8, 37.9, 34.5, 1, 34.5, 1)
F_measure <- c(79.2, 85.2, 79.3, 1, 83.4, 1)
AUC <- c(38.43, 35.85, 62.6, 60.0, 63.81, 1)

ktable1 <- as.matrix(data.frame(Model, Recall, Precision, Accuracy, Sensitivity, Specificity, F_measure, AUC))
ktable1

```

**CREATIVELY PLOT IF THIS IS THE WAY WE GO**
```{r}
#barplot(matrix, col = "lightblue",border = "lightblue" ,
 #        main = "Classification Tree Comparison", ylab = "RMSE", ylim=c(0,4000))
```


```{r}
#plot(rocCurve.tree,col = c(4))
#plot(rocCurve.bagg,add = TRUE, col = c(6)) # color magenta is bagg
#plot(rocCurve.rf,add=TRUE,col=c(1)) # color black is rf
#plot(rocCurve.cf,add=TRUE,col=c(2)) # color red is cforest
#plot(rocCurve.gbm,add=TRUE,col=c(3)) # color green is gbm
```


### HELP!!!
1. Should we add OOB? Some of the models have that feature built within their function, not sure if we have enough already?
This will depend which method we go with (LDA, trees, log, etc.)
2. Need to present the conclusion in an optimal way. We can work on this when we pick the route we will take :)
3. Pruning... make sense?


## Trees: Important Notes

Resource: <https://fderyckel.github.io/machinelearningwithr/trees-and-classification.html> Important Terminology related to Decision Trees

Root Node: It represents entire population or sample and this further gets divided into two or more homogeneous sets.

Splitting: It is a process of dividing a node into two or more sub-nodes.

Decision Node: When a sub-node splits into further sub-nodes, then it is called decision node.

Leaf/ Terminal Node: Nodes do not split is called Leaf or Terminal node.

Pruning: When we remove sub-nodes of a decision node, this process is called pruning. You can say opposite process of splitting.

Branch / Sub-Tree: A sub section of entire tree is called branch or sub-tree.

Parent and Child Node: A node, which is divided into sub-nodes is called parent node of sub-nodes where as sub-nodes are the child of parent node.

9.3, 13.1, 13.2



```{r packages}
library(data.table)
library(ggplot2)
library(e1071)
library(missForest)
library(caret)
library(MASS)
library(gridExtra)
```

```{r}
aids <- fread("Caids.csv")
aids[, sex := as.factor(sex)]
aids[, type := factor(type, levels = c("DP", "CP"), ordered = TRUE)]
```


```{r}
set.seed(256)
aids_completed <- missForest(aids, variablewise = TRUE)
aids <- aids_completed$ximp
```

```{r}
aids[, sex := as.numeric(sex) - 1]
aids[, type := as.numeric(type) - 1]
```

## Trees: Data Partitioning

```{r}
library(rpart) # Used for building classification and regression trees
library(RColorBrewer)
library(rattle)  # Graphical Data Interface
library(rpart.plot) # Automatically scales and adjusts the displayed tree for best fit
library(gains)
library(adabag)
library(randomForest)
library(lift) # TopDecileLift()
library(caret)
library(kableExtra)
library(vip)
library(tidyverse) # Data manipulation
library(xgboost) # Fitting the xgboost model
library(pROC)
library(ipred)       #for fitting bagged decision trees

# Set the seed for the random number generator for reproducing the partition.
set.seed(1)

# Partitioning into training (60%) and validation (40%) 
train.index <- sample(c(1:dim(aids)[1]), dim(aids)[1]*0.6)
test.index <- setdiff(c(1:dim(aids)[1]), train.index)  

# Collect all the columns with training rows into training set 
train <- aids[train.index, ]
test <- aids[test.index, ]
```

## Deep Classification Decision Tree

```{r Deep Tree}
# Set the seed for the random number generator for reproducing the partition.
set.seed(450)

# Create deep tree (page 219)
deep_class_tree <- rpart(formula = type ~ ., data = train, method = "class", control = rpart.control(maxdepth = 30))

# Generates a cost complexity parameter table that provides the complexity parameter value
#summary(deep_class_tree)

# Plot tree
prp(deep_class_tree, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10, box.col  = ifelse(deep_class_tree$frame$var == "<leaf>", 'lightblue', 'white'))

# Count number of leaves 
length(deep_class_tree$frame$var[deep_class_tree$frame$var == "<leaf>"]) 

# Variable importance, deep tree
kable_styling(kable(round(100*deep_class_tree$variable.importance / sum(deep_class_tree$variable.importance), 2), col.names = "Importance %"), full_width = TRUE) |>  kable_classic() 

printcp(deep_class_tree)

#Get the lowest CP value from CP table
deep_class_tree$cptable[which.min(deep_class_tree$cptable[,"xerror"]),"CP"]

predict_model2 <- predict(deep_class_tree, test, type = "class")

# Create confusion matrix for classification tree
conf_mat_tree1  <- confusionMatrix(predict_model2, as.factor(test$type))
conf_mat_tree1

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_model2, as.factor(test$type))$table)



```

```{r}
# Set the seed for the random number generator for reproducing the partition.
set.seed(453)

model = rpart(formula = type ~ ., data = train, method = "class")

rpart.plot(model)
print(model) 

predict_model <- predict(model, test, type = "class")

# Create confusion matrix for classification tree
conf_mat_tree1  <- confusionMatrix(predict_model, as.factor(test$type), positive = "1")
conf_mat_tree1

# Create fourfoldplot
fourfoldplot(confusionMatrix(predict_model,as.factor(test$type))$table, conf.level = 0)

fourfoldplot(confusionMatrix(predict_model,as.factor(test$type))$table, conf.level = 0, std = "all.max", color = c("#f98e5a", "#9cc9de"))

# Variable importance, deep tree
kable_styling(kable(round(100*model$variable.importance / sum(model$variable.importance), 2), col.names = "Importance %"), full_width = TRUE) |>  kable_classic() 

fancyRpartPlot(model, uniform = TRUE, caption = NULL, palettes = "OrRd", digits = -3)


```


```{r}
type_pred <- predict(model, newdata = test, type = 'prob')

type_pred_class <- ifelse(type_pred >= 0.3, 1, 0)

confusionMatrix(data = as.factor(type_pred_class[,2]), reference = as.factor(test$type), positive = "1")
```

```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a data frame with the data and labels
df <- data.frame(data = c(44.43, 31.36, 23.81, 0.40),
                 labels = c("rna", "cd8", "cd4", "sex"))

df$labels <- reorder(df$labels, +df$data)

# Create the plot using ggplot2
ggplot(df, aes(x = data, y = labels, fill = "#9cc9de")) +
  geom_bar(stat = "identity") +
  labs(x = "Percentage", y = "Predictors") + 
  scale_fill_identity() +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10)) 
```

-   Upper left: True Negative - You predicted negative and it's true.\
-   Lower left: False Negative - You predicted negative and it's false, leading to a Type II error\
-   Upper right: False Positive - You predicted positive and it's false, which is a Type I error\
-   Lower right: True Positive - You predicted positive and it's true.\

Recall: From all positive classes, how many we predicted correctly. This value should be as high as possible.\
$$Recall =  \frac{TP}{TP + FN}$$

```{r Deep Tree Recall}
deep_recall1 <- conf_mat_tree1$byClass[6]
deep_recall1
```

The recall score is 78.3%.\

Precision: From all the classes we have predicted as positive, how many are actually positive. This value should be as high as possible.\
$$Recall =  \frac{TP}{TP + FP}$$

```{r Deep Tree Precision}
deep_precision1 <- conf_mat_tree1$byClass[5]
deep_precision1
```

The precision score is 73.8%.\

```{r Acc Spec Spec Deep Tree}
# Accuracy and specificity for tree
tree_res1 = c(conf_mat_tree1$overall[1], conf_mat_tree1$byClass[1], conf_mat_tree1$byClass[2])

tree_res1
```

Accuracy: From all the classes (positive and negative), how many of them we have predicted correctly. This value should be as high as possible.\
Sensitivity: The proportion of true positives among all actual positives. In other words, sensitivity measures how well the model identifies the positive class. Higher the value, the better!\
$$ Sensitivty = TP/(TP+FN)$$

Specificity: The proportion of true negatives among all actual negatives. In other words, specificity measures how well the model identifies the negative class. Higher the value, the better!\
$$ Specificity = TN/(TN+FP)$$

The accuracy s 63.3%, while Sensitivity is 78.3% and Specificity is 20.7%. This model does not predict the negative class as well.\

F-measure: It is difficult to compare two models with low precision and high recall or vice versa. So to make them comparable, we use F-Score. F-score helps to measure Recall and Precision at the same time. It uses Harmonic Mean in place of Arithmetic Mean by punishing the extreme values more.\
$$F-measure =  \frac{2*Recall*Precision}{Recall + Precision}$$

```{r Deep Tree F-measure}
(2*deep_recall1 * deep_precision1)/(deep_recall1 + deep_precision1)
```

The f-measure score is 76.2%.

```{r}
library(ROCR)
# ROC curve to choose cutoff
type_pred <- predict(model, newdata = test, type = 'prob')
type_pred
# prediction() creates an object of "formal class prediction" which contains information about the
# predicted probabilities and the actual values of the validation set.
Predictions = prediction(type_pred[, 2], test$type)
Predictions
# our "formal class prediction" object serves as input for the performance() function.
# the second argument of performance will specify which characteristic should be computed with
# respect to the cutoff value. It computes the characteristic for 2000 different values of the cutoff
SensPerf = performance(Predictions, "sens")  # computing sensitivity wrt cutoff
SensPerf
SpecPerf = performance(Predictions, "spec")  # computing specificity wrt cutoff
SpecPerf
# SensPerf and SpecPerf are "formal class performance" objects.
AccPerf = performance(Predictions, "acc")
AccPerf

```


```{r}
# we create a data table which contains the cutoff, the sensitivity and the specificity values
CutSensSpec = data.table(     Cutoff = SensPerf@x.values[[1]],  # use @ to access content inside
                         Sensitivity = SensPerf@y.values[[1]],  # formal class (S4) objects
                         Specificity = SpecPerf@y.values[[1]],
                            Accuracy =  AccPerf@y.values[[1]])
CutSensSpec[1, 1] = 1
```

```{r}
CutSensSpec <- CutSensSpec[order(nrow(CutSensSpec):1),]
CutSensSpec <- CutSensSpec[rep(seq_len(nrow(CutSensSpec)), each = 2), ]
CutSensSpec[2:15, 2:4] <- CutSensSpec[3:16, 2:4]
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Sensitivity, color = "Sensitivity")) +
  geom_line(aes(y = Specificity, color = "Specificity")) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
       y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .5, linetype = "dashed", color = "black") +scale_y_continuous(limits = c(0,1))
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
              y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .5, linetype = "dashed", color = "black") +scale_y_continuous(limits = c(0,1))
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
              y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .25, linetype = "dashed", color = "black") +scale_y_continuous(limits = c(0,1))
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Sensitivity, color = "Sensitivity")) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
      y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .25, linetype = "dashed", color = "black") +scale_y_continuous(limits = c(0,1))
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Sensitivity, color = "Sensitivity")) +
  geom_line(aes(y = Specificity, color = "Specificity")) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
       y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .25, linetype = "dashed", color = "black") +scale_y_continuous(limits = c(0,1))
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff)) +
  geom_line(aes(y = Sensitivity, color = "Sensitivity")) +
  geom_line(aes(y = Specificity, color = "Specificity")) +
  geom_line(aes(y = Accuracy, color = "Accuracy")) +
  scale_color_manual(name = "", 
                     values = c("Sensitivity" = "#f98e5a", 
                                "Specificity" = "#9cc9de", 
                                "Accuracy" = "blue")) +
  labs(x = "Cutoff Value", 
        y = "Accuracy, Specificity, orSensitivity (fraction)",
       title = "Accuracy, Specificity, and Sensitivity versus Cutoff Value") +
  theme_minimal()+
  geom_vline(xintercept = .25, linetype = "dashed", color = "black") +geom_vline(xintercept = 0.2875143, linetype = "dashed", color = "red") +scale_y_continuous(limits = c(0,1))
```

```{r}
type_pred <- predict(model, newdata = test, type = 'prob')
type_pred_class <- ifelse(type_pred[, 2] >= 0.25, 1, 0)
confusionMatrix(data = as.factor(type_pred_class), reference = as.factor(test$type), positive = "1")
```

```{r}
 fourfoldplot(confusionMatrix(data = as.factor(type_pred_class), reference = as.factor(test$type), positive = "1")$table, conf.level = 0, std = "all.max", color = c("#f98e5a", "#9cc9de"))
```






