---
title: "Case Study C - HIV"
author: "Oumaima Al Qoh, Francisco Arrieta, Lucia Camenisch, Manuela Giansante, Emily Schmidt, Camille Beatrice Valera"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true # creating a table of contents (toc)
    toc_float: 
      collapsed: false # toc does not collapse and is shown as a sidebar (toc_float)
    number_sections: true # document sections are numbered
    theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

```{r packages}
library(data.table)
library(missForest)
library(caret)       # confusion matrix
library(ROCR)        # ROC curve to choose cutoff
library(ggplot2)
library(forecast)    # BoxCox
# library(e1071)
# library(MASS)
# library(gridExtra)
# library(fabletools)
```

# import data

```{r}
aids <- fread("Caids.csv")
aids[, sex := as.factor(sex)]
aids[, type := factor(type, levels = c("DP", "CP"), ordered = TRUE)]
aids[, nas := ifelse(is.na(rna), 1, 0)]
```

# missing values imputation

```{r}
set.seed(256)
aids_completed <- missForest(aids, variablewise = TRUE)
aids <- aids_completed$ximp
```

```{r}
aids[, sex := as.numeric(sex) - 1]
aids[, type := as.numeric(type) - 1]
```

# train-test partitioning

```{r}
# Partitioning into training (60%) and validation (40%) 
set.seed(1)
train.index <- sample(c(1:dim(aids)[1]), dim(aids)[1]*0.6)
test.index <- setdiff(c(1:dim(aids)[1]), train.index)  

# Collect all the columns with training rows into training set 
train <- aids[train.index, ]
test <- aids[test.index, ]

mean(train$nas)
mean(test$nas)
```

# scaling numerical predictors


```{r}
scaled_train <- scale(train[, c(3:5)])
train$cd4 <- scaled_train[, 1]
train$cd8 <- scaled_train[, 2]
train$rna <- scaled_train[, 3]
```


```{r}
means <- attributes(scaled_train)$`scaled:center`
sds <- attributes(scaled_train)$`scaled:scale`
means
sds
```

```{r}
test[, cd4 := (cd4 - means[1])/sds[1]]
test[, cd8 := (cd8 - means[2])/sds[2]]
test[, rna := (rna - means[3])/sds[3]]
```



# basic logistic regression

With cutoff at 0.5, logistic regression with

- cd8, rna
- cd8, rna, sex
- cd8, rna, cd4
- cd8, rna, sex, cd4

all give exactly the same predictions. All models which do not contain cd8 and rna perform worse.


```{r}
model <- glm(formula =  type ~ cd8 + rna, data = train, family = 'binomial')
summary(model)
```

# logistic regression confusion matrix

```{r}
# Predict CD4, CD8, and RNA counts for testing set
type_pred <- predict(model, newdata = test, type = 'response')
type_pred <- round(type_pred)

CM <- confusionMatrix(data = as.factor(type_pred), reference = as.factor(test$type), positive = "1")
CM
```

# ROC curve

```{r}
type_pred <- predict(model, newdata = test, type = 'response')
# prediction() creates an object of "formal class prediction" which contains information about the
# predicted probabilities and the actual values of the validation set.
Predictions = prediction(predictions = type_pred, labels = test$type)
Predictions
# our "formal class prediction" object serves as input for the performance() function.
# the second argument of performance will specify which characteristic should be computed with
# respect to the cutoff value. It computes the characteristic for 2000 different values of the cutoff
SensPerf = performance(Predictions, "sens")  # computing sensitivity wrt cutoff
SensPerf
SpecPerf = performance(Predictions, "spec")  # computing specificity wrt cutoff
SpecPerf
# SensPerf and SpecPerf are "formal class performance" objects.
AccPerf = performance(Predictions, "acc")
AccPerf
```
```{r}
# we create a data table which contains the cutoff, the sensitivity and the specificity values
CutSensSpec = data.table(     Cutoff = SensPerf@x.values[[1]],  # use @ to access content inside
                         Sensitivity = SensPerf@y.values[[1]],  # formal class (S4) objects
                         Specificity = SpecPerf@y.values[[1]],
                            Accuracy =  AccPerf@y.values[[1]])
CutSensSpec[1, 1] = 1
```

```{r}
ggplot(CutSensSpec, aes(x = Cutoff, y = Sensitivity, color = "Sensitivity")) + geom_line() +
  geom_line(mapping = aes(y = Specificity, color = "Specificity")) +
  geom_line(mapping = aes(y = Accuracy, color = "Accuracy")) +
  theme_light() +                                                # aesthetic theme of plot
  labs(x = "Cutoff Value",                                       # labels of the plot
       y = "Specificity or Sensitivity (fraction)",
       title = "Specificity and Sensitivity versus Cutoff Value",
       color = "")
```

```{r}
# CutSensSpec[, Diff := abs(Specificity - Sensitivity)]   # creating Diff
# MinDiff = CutSensSpec[Diff %in% min(Diff),]             # getting the line in table with OptCutoff
# MinDiff
# OptCutoff = as.numeric(MinDiff[1, 1])                   # getting OptCutoff
```

```{r}
type_pred <- predict(model, newdata = test, type = 'response')
type_pred_class <- ifelse(type_pred >= 0.3, 1, 0)
confusionMatrix(data = as.factor(type_pred_class), reference = as.factor(test$type), positive = "1")
```

# linearity of log odds of predictors wrt to predicted values

```{r}
test_log_odds <- log(type_pred / (1 - type_pred))
```


```{r}
ggplot(mapping = aes(x = test$cd8, y = test_log_odds)) + geom_point()
ggplot(mapping = aes(x = test$rna, y = test_log_odds, color = as.factor(test$nas))) + geom_point()

ggplot(mapping = aes(x = train$cd8, y = model$linear.predictors)) + geom_point()
ggplot(mapping = aes(x = train$rna, y = model$linear.predictors, color = as.factor(train$nas))) + geom_point()
```





